Radiative Convective Equilibrium as A Framework for Studying the Interaction Between Convection and its Large-scale Environment

### 2019-09-20 Silvers_etal_2016_sbmt.tex
committing a number of old tex files into the latex_git repo

\begin{abstract}
Simulations of parameterized Radiative Convective Equilibrium (RCE) have been analyzed across a range of domains with areas spanning more than two orders of magnitude ($\mathrm{0.80 - 204 \times 10^6 km^2}$).  It is shown that the simulated atmosphere of a General Circulation Model (GCM) in RCE is sufficiently similar across a range of domain sizes to justify the use of RCE to study both a GCM and a high resolution model on the same domain with the goal of improved constraints on the parameterized clouds.  The simulated climates on different domains are qualitatively similar in their degree of convective organization, the precipitation rates, and the vertical structure of the clouds, relative humidity, and water vapor.  Sea surface temperature perturbation experiments are used to estimate the climate feedback parameter for the differently configured experiments, and the cloud radiative effect is computed to examine the role which clouds play in the response.   Despite the similar climate states between the domains the feedback parameter varies by more than a factor of two; the hydrological sensitivity parameter is better behaved, varying by a factor of 1.4.  The sensitivity to domain size is related foremost to a non-systematic response of low-level clouds as well as an increasingly negative longwave feedbacks on larger domains.  
\end{abstract}

@@

Research over at least the past two decades has shown that an uncertain representation 
of clouds, specifically tropical low-level clouds, is currently the largest source of spread among state-of-the-art climate models [\textit{Cess et al.}, 1989\nocite{Cess_etal_1989}, \nocite{Bony_Dufresne_2005} \textit{Bony and Dufresne}, 2005, \nocite{Stevens_Bony_2013}\textit{Stevens and Bony}, 2013\nocite{Popke_etal_2013}].  Specific limitations of modeled clouds include both features which differ from what one observes (e.g., `too few, too bright', \textit{Webb et al.} [2001]\nocite{Webb_etal_2001}, \textit{Nam et al.} [2012]\nocite{Nam_etal_2012}) as well as features which are inadequately understood
(e.g., controls on cloud vertical structure, \textit{Nuijens et al.} [2015]\nocite{Nuijens_etal_2015}, and organization, \textit{Bretherton et al.} [2005], \textit{Muller and Held} [2012], \textit{Wing and Emanuel} [2014]).
An important reason why clouds remain inadequately represented in models, and poorly understood is the  
%These types of features in General Circulation Model (GCM) simulations hint at the interplay between convective and stratiform clouds, as well as the 
delicacy of the interactions between clouds and circulations (\textit{Bony et al.} [2015]\nocite{Bony_etal_2015}, \textit{Sherwood et al.} [2014]\nocite{Sherwood_etal_2014}). 
%which makes it difficult to separate cause and effect.
The characteristics of clouds in GCMs should be represented
in a manner which is consistent with fundamental theory, observations, and more fundamental models of cloud processes, for instance Cloud Resolving Models (CRMs).
Efforts to better understand the role of moist convection in the climate system would be advanced by an ability to simulate the interaction between convection and the large-scale circulation using both models with parameterized convection and models with explicitly simulated convection.     
However, for these types of problems, making a logical and clear connection between CRMs, and large-scale models that parameterize convection is difficult in large part because of the immensely differing range of parameters and large differences in the range of scales simulated, particularly for climate applications.

Global, or large-scale models using Radiative Convective Equilibrium (RCE) with parameterized convection have led to insights concerning tropical energetics of GCMs and the climate sensitivity (\textit{Popke et al.} [2013], \textit{Becker and Stevens} [2014]\nocite{Becker_Stevens_2014}, \textit{Reed et al.} [2015]\nocite{Reed_etal_2015}), and are increasingly being used to study the interplay between convection and large-scale circulations [\textit{Arnold and Randall}, 2015]\nocite{Arnold_Randall_2015}.  
It has also become clear that even with the relative simplicity of RCE, differences in convective parameterization schemes   
%and other processes dependent determined by subgrid-scale parameterizations, 
play a decisive role on the overall statistics of the simulation [\textit{Popke et al.}, 2013, \textit{Becker et al.}, 2016, \textit{Coppin and Bony}, 2015\nocite{Coppin_Bony_2015}].  
Due to the multiplicity of solutions that previous CRM-RCE simulations [\textit{Held et al.}, 2007, \textit{Jeevanjee and Romps}, 2013\nocite{Jeevanjee_Romps_2013}, \textit{Emanuel et al.}, 2014\nocite{Emanuel_etal_2014}, and \textit{Zhou}, 2015\nocite{Zhou_2015}] have produced which depend on parameterizations, domain size, grid-spacing, and surface temperature, we cannot per se assume that solutions of GCM-RCE simulations will be independent of domain size.  Examining these domain size effects using simulations with parameterized convection, extending to near global scales, is the aim of this paper.

@@

Although not specifically focused on details of the convection or cloud feedbacks, early studies of RCE [\textit{Manabe and Strickler}, 1964\nocite{Manabe_Strickler_1964}, \textit{Manabe and Wetherald}, 1967\nocite{Manabe_Wetherald_1967}] led to a fundamental advancement in our understanding of the atmosphere.  The overall surface warming from high clouds and cooling from low clouds was first shown by Manabe and Strickler using this ansatz.   Manabe and Wetherald showed the equilibrium surface temperature warming in RCE, 
for a specified change in atmospheric CO$_2$,
is strongly dependent on whether atmospheric relative humidity or absolute humidity is assumed fixed as the surface is warmed.  
In the ensuing decades cloud resolving models were applied to the context of RCE, albeit less in the context of climate change.  So doing allowed assumptions about the structure of the atmosphere, convection, and clouds to be relaxed as more atmospheric processes could be resolved by the simulations.  These pioneering studies have come back into focus with the realization that convective aggregation arises spontaneously from simulations with an explicit representation of convection (e.g. \textit{Held et al.}, [1993]\nocite{Held_etal_1993}, \textit{Tompkins and Craig}, [1998]\nocite{Tompkins_Craig_1998}, \nocite{Bretherton_etal_2005}\textit{Bretherton et al.}, [2005], \nocite{Muller_Held_2012}\textit{Muller and Held}, [2012]), in ways that depend on temperature [\textit{Wing and Emanuel}, 2014]\nocite{Wing_Emanuel_2014}), thereby raising the question  as to its importance for the climate system [\textit{Bony et al.}, 2015, \textit{Mauritsen and Stevens}, 2015\nocite{Mauritsen_Stevens_2015}, \textit{Coppin and Bony}, 2015]. 
The previously mentioned many studies of RCE cover an immense range of domains and resolutions, from domain sizes that marginally resolve convection and its large-scale environment to global scales.

@@

The design of the simulations and details of the model used are presented in the next section.  In section 3 results are presented both in terms of the mean atmospheric state, and the response to a perturbation of 4 K to the SST.  Section 3 also 
explores robust differences in the response to warming, and suggests that 
%the framework constrains many aspects of the changes,
many changes in the simulated atmosphere do not depend on domains size, while the climate sensitivity is sensitive to small differences that are not robust across domain sizes. 
Despite having the same core model and identical physics, the total radiative feedback from warming still varies by about a factor of two.   
Discussion and final conclusions are given in section 4.

@@

The framework of RCE is used here as a tool which allows us to focus on the interactions between radiation, large-scale circulation, and clouds without the complicating influence of
heterogeneities imposed on the large scale.   Despite the homogeneity of the forcing, the spontaneous emergence of large-scale overturning circulations, as also emphasized by \textit{Popke et al.}, [2013] makes it possible to capture the interplay of clouds with these circulations, in contrast to (for instance) single column model studies. 
%circulation, land surface, or a changing sea surface.  
% driven by heterogeneities imposed exogenously (?implies B.C.'s are outside system).
Two simulations with constant SST values of 297 K and 301 K are made on five doubly periodic domains of different physical extent but all having identical grid spacing.  To eliminate the heterogeneities 
the Coriolis force is set to zero and the insolation and profiles of trace-gases (e.g., ozone) are likewise made independent of the horizontal grid.
The cosine of the zenith angle is $\pi/4$ and the insolation used is $340.3 \, \rm{W/m}^2$, approximately equal to Earth's global annual mean value.  
There is no diurnal cycle and a surface of water at a fixed temperature defines the lower boundary. 
%Complications due to a dynamic ocean and the required long adjustment time are eliminated by prescribing a fixed value of SST.   
One result of this is that neither the surface nor the top of the atmosphere is expected to have a balanced energy budget.  However, the flux of energy into and out of the atmosphere will be balanced once the energy content of the atmosphere equilibrates with its forcing, which, after the initial adjustment period, remains steady in time except for weather related fluctuations.  
This provides a simple framework for analyzing the behavior of the convection and the sensitivity of the feedback processes %which is simpler than AMIP or CMIP experiments.  
yet retains the full complexity of the physics and dynamical equations.   %Within this framework it is possible to study simulations of RCE on doubly periodic domains which span three orders of magnitude in size.

@@

Because there are fewer grid points on the smaller domains, the degree of variability in the domain mean per time step increases as the domain size decreases.  
This is because a given perturbation will have a larger influence on the statistics of a domain with fewer grid points, relative to a domain with more grid points.  
%This can be seen, for example, in the domain mean values of the total precipitation (Figure \ref{fig:precipvtime}).  
To obtain comparable statistics between the various domains, and take advantage of the computational economy of the smaller domains, the length of simulation 
increases as the domain size decreases.  Details of the simulations examined in this paper are given in Table \ref{exptable}.   The periods used for computing feedback and sensitivity parameters and cloud radiative effects (P1) and domain mean profiles (P2) are given in Table \ref{exptable}.   The increased length of P1 was necessary to obtain robust statistics for the feedback parameters of the different domains.

@@

The horizontal grid used in this study consists of equilateral triangles on a plane with periodic boundary conditions in both dimensions.  The ICON model solves the fully three-dimensional nonhydrostatic and compressible Navier-Stokes equations and is described in detail in \textit{Z\"{a}ngl et al.}, [2015]\nocite{Zaengl_etal_2015}.
All of the simulations described here use triangles with an edge length of 20 km, implying an effective grid spacing of about 13 km (defined as $\sqrt{\rm{cell \, area}}$ in \textit{Z\"{a}ngle et al.}, [2015].  The vertical coordinate is a height-based hybrid coordinate with 60 vertical levels between the surface and the model top at 40 km.  The top of the model uses a Rayleigh damping of the vertical wind following \nocite{Klemp_etal_2008}\textit{Klemp et al.}, [2008] to prevent the reflection of gravity waves from the model top. %starting at 25km.  
  The variables are positioned on a regular-triangular C-grid with a combination of first- and second-order finite difference spatial discretizations.  Time integration is computed with a two time-level predictor-corrector scheme with the physics parameterizations split into slow and fast modes.  Further information on discretization aspects of triangular grids are discussed in \textit{Gassmann} [2011]\nocite{Gassmann_2011} and \textit{Wan et al.}, [2013]\nocite{Wan_etal_2013}. The ongoing development of ICON is a joint project between the Max Planck Institute for Meteorology (MPI-M) and the German Weather Service (Deutsche Wetter Dienst: DWD).  Thus ICON has been developed as a unified modeling framework that functions both as an operational weather forecast model and a global climate model.  
%The German Weather Service implemented ICON as their operational weather forecast model in January of 2015.  
ICON can also be used as a Large Eddy Simulation model capable of explicitly representing canonical cloudy boundary layer flows on grid meshes as fine as 50 m or less [\textit{Dipanker et al.}, 2015]. \nocite{Dipanker_etal_2015}

%here (NWP physics, used operationally at DWD) are derived from a combination of the regional COSMO model (Doms and Sch\"{a}ttler 2004) and the Integrated Forecast System (IFS) model at ECMWF.  
The physics parameterizations used 
in the present study are the NWP physics, so named because they are used operationally by DWD in their global implementation of ICON.
The convective parameterization scheme is based on the Tiedtke mass-flux approach with modifications as discussed by \textit{Bechtold et al.}, [2008].  The radiative transfer scheme is an RRTM scheme similar to that used in the IFS model and identical to that of the ECHAM6.1 model [\textit{Mlawer et al.}, 1997\nocite{Mlawer_etal_1997}, \nocite{Stevens_etal_2013}\textit{Stevens et al.}, 2013].  The microphysics scheme is from the COSMO-EU five-category prognostic scheme [\textit{Seifert}, 2008] predicting rain, snow, cloud ice, cloud liquid water, and water vapor. 
Relative humidity is computed over a liquid surface only.  
The cloud cover scheme is based on a box-function probability density function with an added diagnostic anvil term that uses the detrainment of the convective parameterization.  The boundary layer and turbulence scheme relies on a prognostic equation for turbulent kinetic energy [\textit{Raschendorfer}, 2001].  For these experiments gravity waves are not parameterized.   The development of these schemes is ongoing, those used for this study correspond to revision number 16400.  As an additional option to the NWP physics parameterizations used here, a successor to the ECHAM 6 physics is also being implemented into ICON at MPI-M which will be used for climate studies.

 The profile used for the ozone is representative of the tropics and is the same profile used by \textit{Popke et al.}, [2013].  To trigger the adjustment process perturbations of 0.2 K, white in spatial distribution, were applied to the initial potential temperature field in the lowest four model levels.  The initial values of the vertical, $w,$ and one component, $v$, of the horizontal velocity fields are zero.  In the lowest 2.5 km the horizontal velocity $u$ begins with an exponentially decaying profile ($5 \,\rm{m \, s^{-1}}$ at the surface).  
%which is dissipated to near zero within the first few hours of the experiment.  
For the simulations to behave reasonably it was necessary to specify a minimum surface wind velocity scale used only in the computation of the surface fluxes.   This required adding $0.8 \, \rm{m \, s^{-1}}$  to the default value used in the physics ($0.01 \, \rm{m \, s^{-1}}$).  
%to $0.81 \, \rm{m \, s^{-1}}$ where the second significant digit was retained for convenience. 
The initial profiles for potential temperature,  and water vapor are idealized based upon the best fit to profiles from a large-eddy simulation.  The potential temperature was initialized as $\theta=\theta_{\rm{sfc}}+4.8 z$ below the tropopause and as $\theta=\theta_{\rm{sfc}}+4.8 z_{\rm{tp}}+22.0(z-z_{\rm{tp}})$ at and above the tropopause, where $\theta$ is potential temperature ($\rm{K}$), $z$ is height ($\rm{km}$) and the subscripts `tp' and `sfc' represent the values at the tropopause level and surface, respectively.   The water vapor was initialized as an exponentially decaying concentration with a scale height of 2.8 km.
%To facilitate the proper coupling between the surface and the atmosphere under low-wind conditions the minimum value of the %wind vector which is used in the computation of the bulk surface flux formulae was changed from 0.01 $\rm{m \, s^{-1}}$ to 0.81 %$\rm{m \, s^{-1}}$.  
All experiments use the same grid spacing and time step as discussed previously. 
Our choice of grid cell size with an effective grid spacing of 13 km matches that currently used by DWD operationally.

\section{Results}

To give an impression of the relative scale of features present on the
$3/4 \rm{M}$, $12 \rm{M}$, and $200 \rm{M}$ domains Figure \ref{fig:convorg} shows the same data from Figure \ref{fig:precvsdom}, tiled according to scale, periodically repeated to fill a region the size of the $200 \rm{M}$ domain.  The $3/4 \rm{M}$, $3 \rm{M}$, $12 \rm{M}$, and $50 \rm{M}$ domains all show the precipitation and moisture structures to organize on the largest scale available (Figures \ref{fig:precvsdom}, \ref{fig:convorg}, \ref{fig:tqv5dom}).  The 200M domain shows signs that the maximum scale of convective organization may be between the 50M and 200M domains, but this deserves further investigation.

\subsection{Mean Climate}

One method of diagnosing whether the atmosphere has reached a statistically stationary state is to examine the time evolution of various quantities.  After an initial adjustment period with large fluctuations in the energetics of the simulations  stationarity is approached.  
%Once this state has been reached the prognostic variables oscillate about a constant mean value.  
In stationarity short temporal fluctuations of domain means, which diminish with domain size and time-scale, are still apparent, reflecting the dynamic state of the underlying system.
The evolution of the domain mean total precipitation during the first two years of simulation for each of the five domains is shown in Figure \ref{fig:precipvtime}.  Based on the precipitation we judge that all simulations have
reached a stationary state after about 200 days, with the $50\rm{M}$ and $200\rm{M}$ domains adjusting to equilibrium much sooner.  Because the $3/4\rm{M}$, $3\rm{M}$, and $12\rm{M}$ domains all proceed for longer than 2 years, the average value of the precipitation during the last six months of each simulation is shown by the dots at the right of Figure \ref{fig:precipvtime}, confirming that the mean value has indeed been reached during the period shown.  
%There is a systematic variation  of the evolution of the precipitation field as a function of domain size but the differences are relatively small (10\%) as compared to the mean.
%have differences between the models, overall the similarity is more striking than the difference with a range of only about $10\%$ amongst the models.  The domain mean precipitation increases monotonically with domain size.  This is likely related to the near monotonic 
Variations in the precipitation field are progressively smaller as the model domain size increases with an apparent convergence in the mean value for the largest domains.  The mean (of all domains, with SST$=301 \, \rm{K}$) precipitation is $5 \, \rm{mm \,day^{-1}}$ with no simulation deviating by more than 7 \%.
The systematic increase in precipitation with domain size is consistent with the systematic
decrease of upper level clouds (Figure 5).  A decrease in upper level clouds will increase the atmospheric radiative cooling and thus require more convective heating 
%to maintain balance.    
in stationarity.

>>>>

\begin{abstract}
Simulations of parameterized Radiative Convective Equilibrium (RCE) have been analyzed across a range of domains which span three orders of magnitude.  It is shown that the simulated atmosphere of a General Circulation Model (GCM) in RCE is sufficiently similar across a range of domain sizes to justify the use of RCE to study both a GCM and a high resolution model on the same domain with the goal of improved constraints on the parameterized clouds.  The simulated climates on different domains are qualitatively similar in their degree of convective organization, the precipitation rates, and the vertical structure of the clouds, relative humidity, and water vapor.  Sea surface temperature perturbation experiments are used to estimate the climate feedback parameter for the differently configured experiments, and the cloud radiative effect is computed to examine the role which clouds play in the response.   Despite the similar climate states between the models the feedback parameter varies by more than a factor of two; the hydrological sensitivity parameter is better behaved, varying by a factor of 1.4.  The sensitivity to domain size is related foremost to a non-systematic response of low-level clouds as well as an increasingly negative longwave feedbacks on larger domains.  
\end{abstract}

@@

Research over at least the past two decades has shown that an uncertain representation 
of clouds (specifically tropical low-level clouds) is currently the largest source of spread among state-of-the-art climate models [\textit{Cess et al.}, 1989\nocite{Cess_etal_1989}, \nocite{Bony_Dufresne_2005} \textit{Bony and Dufresne}, 2005, \nocite{Stevens_Bony_2013}\textit{Stevens and Bony}, 2013\nocite{Popke_etal_2013}].  Specific limitations of modeled clouds include both features which differ from what one observes (ex: `too few, too bright', \textit{Webb et al.} [2001]\nocite{Webb_etal_2001}, \textit{Nam et al.} [2012]\nocite{Nam_etal_2012}) as well as features which are inadequately understood
(e.g., controls on cloud vertical structure, \textit{Nuijens et al.} [2015]\nocite{Nuijens_etal_2015}, and organization, e.g., \textit{Bretherton et al.} [2005], \textit{Muller and Held} [2012], \textit{Wing and Emanuel} [2014]).
An important reason why clouds remain inadequately represented in models, and poorly understood is the  
%These types of features in General Circulation Model (GCM) simulations hint at the interplay between convective and stratiform clouds, as well as the 
delicacy of the cloud-circulation interaction (\textit{Bony et al.} [2015]\nocite{Bony_etal_2015}, \textit{Sherwood et al.} [2014]\nocite{Sherwood_etal_2014}) which makes it difficult to separate cause and effect.
The characteristics of clouds in GCMs should be represented
in a manner which is consistent with theory, observations, and more fundamental models of cloud processes, for instance Cloud Resolving Models (CRMs).
Efforts to better understand the role of moist convection in the climate system would be advanced by an ability to simulate the interaction between convection and the large-scale circulation using both models with parameterized convection and models with explicitly simulated convection.     
However, for these types of problems, making a logical and clear connection between CRMs, and large-scale models that parameterize convection is difficult in large part because of the immensely differing range of parameters and large differences in the range of scales simulated, particularly for climate applications.

Global, or large-scale models using Radiative Convective Equilibrium (RCE) with parameterized convection have led to insights concerning tropical energetics of GCMs and the climate sensitivity (\textit{Popke et al.} [2013], \textit{Becker and Stevens} [2014]\nocite{Becker_Stevens_2014}, \textit{Reed et al.} [2015]\nocite{Reed_etal_2015}), and are increasingly being used to study the interplay between convection and large-scale circulations [\textit{Arnold and Randall}, 2015]\nocite{Arnold_Randall_2015}.  
It has also become clear that even with the relative simplicity of RCE, differences in convective parameterization schemes   
%and other processes dependent determined by subgrid-scale parameterizations, 
play a decisive role on the overall statistics of the simulation [\textit{Popke et al.}, 2013, \textit{Becker et al.}, 2016, \textit{Coppin and Bony}, 2015\nocite{Coppin_Stevens_2015}].  
Due to the multiplicity of solutions that previous CRM-RCE simulations [\textit{Held et al.}, 2007, \textit{Jeevanjee and Romps}, 2013\nocite{Jeevanjee_Romps_2013}, \textit{Emanuel et al.}, 2014\nocite{Emanuel_etal_2014}, and \textit{Zhou}, 2015\nocite{Zhou_2015}] have produced which depend on parameterizations, domain size, grid-spacing, and surface temperature, we cannot per se assume that solutions of GCM-RCE simulations will be independent of domain size.  Examining these domain size effects using simulations with parameterized convection, extending to near global scales, is the aim of this paper.

@@

Although not specifically focused on developingtails of the convection or cloud feedbacks, early studies of RCE [\textit{Manabe and Strickler}, 1964\nocite{Manabe_Strickler_1964}, \textit{Manabe and Wetherald}, 1967\nocite{Manabe_Wetherald_1967}] led to a fundamental advancement in our understanding of the atmosphere.  The overall surface warming from high clouds and cooling from low clouds was first shown by Manabe and Strickler using this ansatz.   Manabe and Wetherald showed the equilibrium surface temperature warming in RCE, 
for a specified change in atmospheric CO$_2$,
is strongly dependent on whether atmospheric relative humidity or absolute humidity is assumed fixed as the surface is warmed.  
In the ensuing decades cloud resolving models were applied to the context of RCE, albeit less in the context of climate change.  So doing allowed assumptions about the structure of the atmosphere, convection, and clouds to be relaxed as more atmospheric processes could be resolved by the simulations.  These pioneering studies have come back into focus with the realization that convective aggregation arises spontaneously from simulations with an explicit representation of convection (e.g. \textit{Held et al.}, [1993]\nocite{Held_etal_1993}, \textit{Tompkins and Craig}, [1998]\nocite{Tompkins_Craig_1998}, \nocite{Bretherton_etal_2005}\textit{Bretherton et al.}, [2005], \nocite{Muller_Held_2012}\textit{Muller and Held}, [2012]), in ways that depend on temperature [\textit{Wing and Emanuel}, 2014]\nocite{Wing_Emanuel_2014}), thereby raising the question  as to its importance for the climate system [\textit{Bony et al.}, 2015, \textit{Mauritsen and Stevens}, 2015\nocite{Mauritsen_Stevens_2015}, \textit{Coppin and Bony}, 2015]. 
The previously mentioned many studies of RCE cover an immense range of domains and resolutions, from domain sizes that marginally resolve convection and its large-scale environment to global scales.

@@

The design of the simulations and details of the model used are presented in the next section.  In section three the results are presented both in terms of the mean atmospheric state, and the response to a perturbation of 4 K to the SST.  Section three also 
explores robust differences in the response to warming, and suggests that 
%the framework constrains many aspects of the changes,
many changes in the simulated atmosphere do not depend on domains size, while the climate sensitivity is sensitive to small differences that are not robust across domain sizes. 
Despite having the same core model and identical physics, the total radiative feedback from warming still varies by about a factor of two.   
Discussion and final conclusions are given in sections four.

@@

The framework of RCE is used here as a tool which allows us to focus on the interactions between radiation, large-scale circulation, and clouds without the complicating influence 
heterogeneities imposed on the large scale.   Despite the homogeneity of the forcing, the spontaneous emergence of large-scale overturning circulations, as also emphasized by \textit{Popke et al.}, [2013] makes it possible to capture the interplay of clouds with these circulations, in contrast to (for instance) single column model studies. 
%circulation, land surface, or a changing sea surface.  
% driven by heterogeneities imposed exogenously (?implies B.C.'s are outside system).
Two simulations with constant SST values of 297 K and 301 K are made on five doubly periodic domains of different physical extent but all having identical grid spacing.  To eliminate the heterogeneities 
the Coriolis force is set to zero and the insolation and profiles of trace-gases (e.g., ozone) are likewise made independent of the horizontal grid.   There is no diurnal cycle 
and a surface whose properties are set to those of water at a fixed temperature defines the lower boundary. 
Complications due to a dynamic ocean and the required long adjustment time are eliminated by prescribing a fixed value of SST.   One result of this is that neither the surface nor the top of the atmosphere is expected to have a balanced energy budget.  However, the flux of energy into and out of the atmosphere will be balanced once the energy content of the atmosphere equilibrates with its forcing, which, after the initial adjustment period, remains steady in time except for weather related fluctuations.  
This provides a simple framework for analyzing the behavior of the convection and the sensitivity of the feedback processes %which is simpler than AMIP or CMIP experiments.  
yet retains the full complexity of the physics and dynamical equations.   %Within this framework it is possible to study simulations of RCE on doubly periodic domains which span three orders of magnitude in size.

@@

Because there are fewer grid points on the smaller domains, the degree of variability in the domain mean per time step increases as the domain size decreases.  
%This can be seen, for example, in the domain mean values of the total precipitation (Figure \ref{fig:precipvtime}).  
To obtain comparable statistics between the various domains, and take advantage of the computational economy of the smaller domains, the length of simulation and averaging period increases as the domain size decreases.  
To give an impression of the relative scale of features present on the
$3/4 \rm{M}$, $12 \rm{M}$, and $200 \rm{M}$ domains Figure \ref{fig:convorg} shows them tiled according to scale periodically repeated to fill a region the size of the $200 \rm{M}$ domain.  In general, precipitation is seen to organize on the largest scale available.  The 200M domain shows signs that the maximum scale of convective organization may be between the 50M and 200M domains, but this is not conclusive and deserves further investigation.

@@

The horizontal grid used in this study consists of equilateral triangles on a plane with periodic boundary conditions in both dimensions.  The ICON model solves the fully three-dimensional nonhydrostatic and compressible Navier-Stokes equations and is described in detail in \textit{Z\"{a}ngl et al.}, [2015]\nocite{Zaengl_etal_2015}.  The variables are positioned on a regular-triangular C-grid with a combination of first- and second-order finite difference spatial discretizations.  Time integration is computed with a two time-level predictor-corrector scheme with the physics parameterizations split into slow and fast modes.  Further information on discretization aspects of triangular grids are discussed in \textit{Gassmann} [2011]\nocite{Gassmann_2011} and \textit{Wan et al.}, [2013]\nocite{Wan_etal_2013}. The ongoing development of ICON is a joint project between the Max Planck Institute for Meteorology (MPI-M) and the German Weather Service (Deutsche Wetter Dienst: DWD).  Thus ICON has been developed as a unified modeling framework that functions both as an operational weather forecast model and a global climate model.  
%The German Weather Service implemented ICON as their operational weather forecast model in January of 2015.  
ICON can also be used as a Large Eddy Simulation model capable of explicitly representing cannonical cloudy boundary layer flows on grid meshes as fine as 50 m or less [\textit{Dipanker et al.}, 2015]. \nocite{Dipanker_etal_2015}

%here (NWP physics, used operationally at DWD) are derived from a combination of the regional COSMO model (Doms and Sch\"{a}ttler 2004) and the Integrated Forecast System (IFS) model at ECMWF.  
The physics parameterizations used 
in the present study are the NWP physics, so named because they are used operationally by DWD in their global implementation of ICON.
The convective parameterization scheme is based on the Tiedtke mass-flux approach with modifications as discussed by \textit{Bechtold et al.}, [2008].  The radiative transfer scheme is an RRTM scheme similar to that used in the IFS model and identical to that of the ECHAM6.1 model [\textit{Mlawer et al.}, 1997\nocite{Mlawer_etal_1997}, \nocite{Stevens_etal_2013}\textit{Stevens et al.}, 2013].  The microphysics scheme is from the COSMO-EU five-category prognostic scheme [\textit{Seifert}, 2008] predicting rain, snow, cloud ice, cloud liquid water, and water vapor. The cloud cover scheme is based on a box-function probability density function with an added diagnostic anvil term that uses the detrainment of the convective parameterization.  The boundary layer and turbulence scheme relies on a prognostic equation for turbulent kinetic energy [\textit{Raschendorfer}, 2001].  For these experiments gravity waves are not parameterized.   The development of these schemes is ongoing, those used for this study correspond to revision number 16400.  As an additional option to the NWP physics parameterizations used here, a successor to the ECHAM 6 physics is also being implemented into ICON at MPI-M which will be used for climate studies.

All of the simulations described here use triangles with an edge length of 20 km, implying an effective grid spacing of about 13 km (defined as $\sqrt{\rm{cell \, area}}$ in \textit{Z\"{a}ngle et al.}, [2015], and a model top at 40 km.  The vertical coordinate is a z-based hybrid coordinate with 60 vertical levels, which over a water surface, as used here, default to z-levels.  The top of the model uses a Rayleigh damping of the vertical wind following \nocite{Klemp_etal_2008}\textit{Klemp et al.}, [2008] to prevent the reflection of gravity waves from the model top. %starting at 25km.

Details of the simulations examined in this paper are given in Table \ref{exptable}.  The profile used for the ozone is representative of the tropics and is the same profile used by \textit{Popke et al.}, [2013].  To trigger the adjustment process perturbations of 0.2 K, white in spatial distribution, were applied to the initial potential temperature field in the lowest four model levels.  The initial values of the vertical, $w,$ and one component, $v$, of the horizontal velocity fields are zero.  In the lowest 2.5 km the horizontal velocity $u$ begins with an exponentially decaying profile ($5 \,\rm{m \, s^{-1}}$ at the surface).  
%which is dissipated to near zero within the first few hours of the experiment.  
For the simulations to behave reasonably it was necessary to specify a minimum surface wind velocity scale used only in the computation of the surface fluxes.   This required adding $0.8 \, \rm{m \, s^{-1}}$  to the default value used in the physics ($0.01 \, \rm{m \, s^{-1}}$).  
%to $0.81 \, \rm{m \, s^{-1}}$ where the second significant digit was retained for convenience. 
The initial profiles for potential temperature,  and water vapor are idealized based upon the best fit to profiles from a large-eddy simulation.  The potential temperature was initialized as $\theta=\theta_{\rm{sfc}}+4.8 z$ below the tropopause and as $\theta=\theta_{\rm{sfc}}+4.8 z_{\rm{tp}}+22.0(z-z_{\rm{tp}})$ at and above the tropopause, where $\theta$ is potential temperature ($\rm{K}$), $z$ is height ($\rm{km}$) and the subscripts `tp' and `sfc' represent the values at the tropopause level and surface, respectively.   The water vapor was initialized as an exponentially decaying concentration with a scale height of 2.8 km.
%To facilitate the proper coupling between the surface and the atmosphere under low-wind conditions the minimum value of the %wind vector which is used in the computation of the bulk surface flux formulae was changed from 0.01 $\rm{m \, s^{-1}}$ to 0.81 %$\rm{m \, s^{-1}}$.  
All experiments use the same grid spacing and time step as discussed previously. 
Our choice of grid cell size with an effective grid spacing of 13 km matches that currently used by DWD operationally.

\section{Results}
\subsection{Mean Climate}

One method of diagnosing whether the atmosphere has reached a statistically stationary state is to examine the time evolution of various quantities.  After an initial adjustment period with large fluctuations in the energetics of the simulations  stationarity is approached.  
%Once this state has been reached the prognostic variables oscillate about a constant mean value.  
In stationarity short temporal fluctuations of domain means, which diminish with domain size and time-scale, are still apparent, reflecting the dynamic state of the underlying system.
The evolution of the domain mean total precipitation during the first two years of simulation for each of the five domains is shown in Figure \ref{fig:precipvtime}.  Based on the precipitation we judge the simulations to
have reached a stationary state after about 200 days.  Because the $3/4\rm{M}$, $3\rm{M}$, and $12\rm{M}$ models all proceed for longer than 2 years, the average value of the precipitation during the last six months of each simulation is shown by the dots at the right of Figure \ref{fig:precipvtime}, confirming that the mean value has indeed been reached during the period shown.  
%There is a systematic variation  of the evolution of the precipitation field as a function of domain size but the differences are relatively small (10\%) as compared to the mean.
%have differences between the models, overall the similarity is more striking than the difference with a range of only about $10\%$ amongst the models.  The domain mean precipitation increases monotonically with domain size.  This is likely related to the near monotonic 
Variations in the precipitation field are progressively smaller as the model domain size increases with an apparent convergence in the mean value for the largest domains.  The mean (of all domains, with SST$=301 \, \rm{K}$) precipitation is $5 \, \rm{mm \,day^{-1}}$ with no simulation deviating by more than 7 \%.
The systematic increase in precipitation with domain size is consistent with the systematic
decrease of upper level clouds (Figure 5).  A decrease in upper level clouds will increase the atmospheric radiative cooling and thus require more convective heating 
%to maintain balance.    
in stationarity.

Basic physical characteristics of the simulated atmospheres can be determined by looking at the domain mean 
thermodynamic profiles.
These are shown for the simulations with $\mathrm{SST}=297 \; \mathrm{K}$ in Figure \ref{fig:prof297}.     Cloud liquid water, cloud cover, and RH, all maximize 
in the lowest kilometer.  The RH drops to 
its tropospheric minimum value ($\approx 15\%$) at about 7km.  Between roughly 12km and 16km there is a maximum 
of relative humidity, cloud cover, and cloud ice, due to radiatively driven mass convergence in the upper troposphere in 
clear skies that is satisfied by the detrainment from convective clouds (e.g., \textit{Hartmann and Larson}[2002]\nocite{Hartmann_Larson_2002}, \textit{Zelinka and 
Hartmann} [2011]\nocite{Zelinka_Hartmann_2011}).   All domains have three peaks in the cloud cover (0.6km, 5km, 13km), with the upper two cloud maxima corresponding to anvil detrainment and melting level clouds, respectively.  The simulations also uniformly produce two peaks of cloud liquid water below 4km.  Overall, the profiles plausibly represent the mean state of the tropical atmosphere (over ocean)
and show agreement among all domain sizes on the vertical structure of the atmosphere.
